---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# openalex

<!-- badges: start -->
[![R-CMD-check](https://github.com/KTH-Library/openalex/workflows/R-CMD-check/badge.svg)](https://github.com/KTH-Library/openalex/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of `openalex` is to provide access to data from [OpenAlex](https://openalex.org) - an open and comprehensive catalog of scholarly papers, authors, institutions and more ... - to R through the [Open Alex REST API](https://docs.openalex.org/api)...

## Installation

You can install the current version of `openalex` from [GitHub](https://github.com/kth-library/openalex) with:

``` r
#install.packages("devtools")
devtools::install_github("kth-library/openalex", dependencies = TRUE)
```

## Example

This is a basic example which shows you how to get information for papers and authors:

```{r example, eval=FALSE}

library(openalex)
library(dplyr)
suppressPackageStartupMessages(library(purrr))
library(knitr)

data <- 
  openalex_crawl("works", verbose = TRUE,
    query = openalex:::openalex_query(filter = 
      "institutions.id:I86987016,publication_year:2022"))

long <- openalex_flatten_long(data)

# number of ids
#long %>% filter(name == "id") %>% nrow()
```

## Rate limits

By providing an email adress you enter the "polite pool" which provides even less of rate limiting for API requests.

Either provide it in `~/.Renviron` or use helper fcn `openalex_polite()` to set or unset email:

```{r polite}
library(openalex)

# set an email to use for the session

openalex_polite("you@example.com")

# unset, and use default user agent string...

openalex_polite("")

```

## Data source attribution

When data from `openalex` is displayed publicly, this attribution also needs to be displayed:

```{r attribution}
library(openalex)
openalex_attribution()
```
